\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\title{Submanifold Geometry}
\author{Paul Bryan}

\begin{document}

\maketitle

\section{Isometric Sub-manifold Geometry}

Let $F: (M, \bar{g}) \to (N, \hat{g})$ be an isometric immersion. Let $\bar{\nabla}$ and $\hat{\nabla}$ denote the respective Levi-Civita connections. On the pull-back bundle $F^{\ast} TN$, we have an induced metric $\hat{g}$ and metric-compatible connection $\hat{\nabla}$. The metric $\hat{g}$ splits the  pull-back bundle into orthogonal subspaces,
\[
F^{\ast}TN \simeq TM \oplus \nu M.
\]
Here we abuse notation, but writing $TM$ for $F_{\ast} TM$ since if $F$ is an immersion, $F_{\ast} : TM \to F^{\ast} TN$ is an injective bundle map and hence isomorphic with it's image. Let $\nu \in \Gamma(F^{\ast}TN)$ be a smooth, unit normal vector field along $M$, i.e. $\hat{g}(\nu, \nu) \equiv 1$ and for any $X \in TM$, $\hat{g}(X, \nu) = 0$.

Let $X, Y \in \Gamma(TM) \subset \Gamma(F^{\ast}TN)$, and define the second fundamental form, $\bar{h} \in \Gamma(T^{\ast} M \otimes T^{\ast} M \otimes \nu M)$,
\[
\bar{h}(X, Y) = \pi^{\nu M} \hat{\nabla}_X Y = \hat{\nabla}_X Y - \pi^{TM} \hat{\nabla}_X Y = \hat{\nabla}_X Y - \bar{\nabla}_X Y,
\]
by the Gauss equation: $\bar{\nabla}_X Y = \pi^{TM} \hat{\nabla}_X Y$. We may thus write
\[
\hat{\nabla}_X Y = \bar{\nabla}_X Y + h(X, Y),
\]
with respect to the splitting $F^{\ast}TN = TM \oplus \nu M$. Observer that $\bar{h}$ is symmetric,
\[
\begin{split}
\bar{h} (X, Y) - \bar{h} (Y, X) &= \hat{\nabla}_X Y - \bar{\nabla}_X Y - \hat{\nabla}_Y X + \bar{\nabla}_Y X \\
&= \hat{\nabla}_X Y - \hat{\nabla}_Y X + \bar{\nabla}_Y X - \bar{\nabla}_X Y \\
&= [X, Y] + [Y, X] = 0
\end{split}
\]
since both connections $\hat{\nabla}$ and $\bar{\nabla}$ are symmetric, and because $dF [X, Y] = [dF X, dF Y]$ so that the Lie bracket is invariant under our identification $dF : TM \simeq dF(TM)$.

Also define, $A \in \Gamma(T^{\ast}M \otimes T^{\ast} M)$,
\[
\bar{A}(X, Y) = \hat{g}(\bar{h}(X, Y), \nu) = \bar{g} (\hat{\nabla}_X Y, \nu),
\]
the second fundamental form with respect to $\nu$, so that,
\[
\bar{h}(X, Y) = \bar{A}(X, Y) \nu,
\]
and $\bar{A}$ is symmetric: $\bar{A}(X, Y) = \bar{A}(Y, X)$.

Define the Weingarten map with respect to $\nu$, $W \in \Gamma(T^{\ast}M \otimes \nu M)$ by
\[
\hat{W}(X) = - \hat{\nabla}_X \nu.
\]

Since $\hat{g}(Y, \nu) = 0$ for any tangent field, we have
\[
0 = \hat{\nabla}_X \hat{g} (Y, \nu) = \hat{g}(\hat{\nabla}_X Y, \nu) + \hat{g} (Y, \hat{\nabla}_X \nu) = \bar{A}(X, Y) - \hat{g}(Y, \hat{W}(X)).
\]
Thus we find that,
\[
\bar{g} (\hat{W}(X), Y) = \hat{g} (\hat{W}(X), Y) = A(X, Y) = A(Y, X) = \hat{g}(\hat{W}(Y)) = \bar{g}(X, \hat{W}(Y)),
\]
and so $W$ is self adjoint with respect to $\bar{g}$.

Observe that we furnished the Wiengarten map with a caret ($\hat{W}$) rather than with a bar since it's definition is independent of the metric $\bar{g}$. However, as we have just seen, since $F$ is isometric, $W$ is self-adjoint with respect to $\bar{g}$ through it's relation to the second fundamental form. 

\section{Non Isometric Submanifold Geometry}

We retain all the notation of the previous section. Now let $g$ be a metric on $M$; in general $g \ne \bar{g}$. Let $\nabla$ be the Levi-Civita connection with respect to $g$. Define, $h \in \Gamma(T^{\ast} M \otimes T^{\ast}M \otimes F^{\ast}TN$,
\[
h(X, Y) = \hat{\nabla}_X Y - \nabla_X Y.
\]
As in the previous section, we may write
\[
\hat{\nabla}_X Y = \nabla_X Y + h (X, Y),
\]
but this is no longer with respect to the splitting $F^{\ast}TN = TM \oplus \nu M$ since $\nabla_X Y \ne \pi^{TM} \hat{\nabla}_X Y$ in general. It is still true however, that $h$ is symmetric since $\hat{\nabla}$ and $\nabla$ are symmetric.

Now it is not so clear how we should define $A$: since $h$ will not be necessarily be normal bundle valued, it doesn't make a lot of sense to define $A(X, Y) = \hat{g}(h(X, Y), \nu)$. Instead, we work with $\hat{W}$. Define $A \in T^{\ast}M \otimes T^{\ast}M)$,
\[
A (X, Y) = g(\hat{W}(X), Y) = - g(\hat{\nabla}_X \nu, Y).
\]
Note that $A(X, Y) \ne \hat{g} (X, Y)$ since $g \ne \bar{g}$ and so we cannot use metric compatibility of $\hat{\nabla}$ with $\hat{g}$ to conclude that $\hat{W}$ is self-adjoint with respect to $g$. Thus we find an incompatibility between $h$ and $A$ (or equivalently between $h$ and $\hat{W}$). Note that $A$ and $\hat{W}$ are however related, the one being a metric contraction with respect to $g$ of the other.

Let us introduce new tensors to measure the incompatibility of $g$ and $\hat{g}$ since this is the source of incompatibilities above. Define,
\[
\beta(X, Y) = g(X, Y) - \hat{g}(X, Y).
\]
The analogous definition for $\bar{g}$, $\bar{\beta}(X, Y) = \bar{g}(X, Y) - \hat{g}(X, Y) = 0$.

\end{document}
